\section{Auxiliary Demonstration: Grammar Parsability vs. Production Validation}
\label{appendix:antlr_vs_syside}

To support the design choice of using production validation (rather than grammar parsing alone) as the acceptance oracle, we ran an auxiliary demonstration included in the repository under \texttt{experiments/antlr\_vs\_syside/}~\cite{antlrVsSysideDemo2026}. This demonstration is not a second primary experiment; it is a construct-validity check showing that parser acceptance and production acceptance are distinct outcomes.

We evaluated 10 intentionally distinct SysMLv2 examples in \texttt{examples/mismatch\_10\_distinct/}. Each file was designed to remain grammar-parseable while violating a model-wide constraint typically enforced by a production validator. The evaluation pipeline was:
\begin{enumerate}
    \item ANTLR parse check (third-party SysML parser integration in the repository).
    \item Production validation check using SysIDE.
\end{enumerate}

Results were unambiguous: all 10/10 examples passed ANTLR parsing, while 0/10 were accepted by production validation (10/10 mismatch cases). Validator diagnostics were dominated by unresolved-reference failures (9/10, \texttt{reference-error}), with one invocation-typing failure (1/10, \texttt{invocation-expression-instantiated-type}). This pattern directly illustrates that context-free syntax conformance is necessary but not sufficient for operational model acceptance in production modeling environments.

\subsection*{Concrete Example}
The following example is grammar-parseable but fails production validation:
\begin{quote}
\small\ttfamily
package Demo07 \{\\
\ \ part def Wheel \{\\
\ \ \ \ attribute hubDiameter: LengthValue;\\
\ \ \ \ part tire \{ attribute width: LengthValue; \}\\
\ \ \ \ attribute outerDiameter: LengthValue = (hubDiameter + 2 * tire.height);\\
\ \ \}\\
\}
\end{quote}

ANTLR parsing accepts the structure. Production validation rejects it with:
\begin{quote}
\small\ttfamily
error (reference-error): No Feature named 'height' found.
\end{quote}

\begin{table}[H]
\centering
\caption{Ten-case auxiliary demonstration: all examples parse under ANTLR, none pass production validation.}
\label{tab:antlr_vs_syside_10cases}
\scriptsize
\setlength{\tabcolsep}{3pt}
\begin{tabular}{p{0.09\columnwidth} p{0.51\columnwidth} p{0.14\columnwidth} p{0.20\columnwidth}}
\toprule
\textbf{ID} & \textbf{Injected Condition} & \textbf{ANTLR} & \textbf{SysIDE} \\
\midrule
01 & Missing imported namespace & Pass & Fail (reference) \\
02 & Missing port type & Pass & Fail (reference) \\
03 & Missing attribute type & Pass & Fail (reference) \\
04 & Missing specialization base & Pass & Fail (reference) \\
05 & Action typed by undefined behavior type & Pass & Fail (reference) \\
06 & Invocation target is not a behavior & Pass & Fail (invocation) \\
07 & Missing referenced feature in expression & Pass & Fail (reference) \\
08 & Missing root-qualified namespace & Pass & Fail (reference) \\
09 & Redefinition of missing feature & Pass & Fail (reference) \\
10 & Missing namespace in type use & Pass & Fail (reference) \\
\bottomrule
\end{tabular}
\end{table}
